---
title: "Temporal_Models"
author: "RC"
date: "11/22/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(geepack)
library(splines2)
library(SimDesign)
library(pracma)
library(mgcv)
library(car)
library(corrplot)
source("getPvalues.r")
```

## Get list of Master data files
Tell which file to look at by setting the value of "i"

```{r find files}
modelDFDir = 'I:/TimeSeries_ScaledByEffortError'
int = '5minBin'
dfList = list.files(path=modelDFDir,pattern=paste('*Master.csv',sep=""),
                    full.names=TRUE,recursive=FALSE,
                    include.dirs=FALSE,no..=TRUE)
i = 1
thisSite = data.frame(read.csv(dfList[i]))
CTname = str_remove(dfList[i],paste(modelDFDir,'/',sep="")) # get the species/CT name
CTname = str_remove(CTname,paste('_',int,'_Master.csv',sep=""))
```

## Assess correlation & collinearity btwn covars 
(adapted from Benjamins '17)


```{r correlation & collinearity}
corrDF = thisSite[,-c(1,2)]
corrDF$DayPhase[which(corrDF$DayPhase=="Dawn")] = 1
corrDF$DayPhase[which(corrDF$DayPhase=="Day")] = 2
corrDF$DayPhase[which(corrDF$DayPhase=="Dusk")] = 3
corrDF$DayPhase[which(corrDF$DayPhase=="Night")] = 4
corrDF$DayPhase = as.numeric(corrDF$DayPhase)
C = cor(corrDF)

corrplot(C,method="ellipse",type="lower",diag=FALSE)

glm1 = glm(Presence~JulianDay+NormTime+LunarIllum+as.factor(Year)+as.factor(DayPhase)
           +as.factor(HATSite),family=binomial,data=thisSite)
vif(glm1)

```

## Compare smooths & linears to null model for continuous covars 
(from Benjamins '17)

``` {r smooth vs lin}

# Julian Day
mod0 = geeglm(Presence~1,family=binomial,data=thisSite,id=GroupID,corstr="ar1")
mod0a = geeglm(Presence~JulianDay,family=binomial,data=thisSite,id=GroupID,corstr="ar1")
mod0b = geeglm(Presence~mSpline(JulianDay,
                                knots=quantile(thisSite$JulianDay, probs=c(0.333,0.666)),
                                Boundary.knots=c(1,365),
                                periodic=T),
               family=binomial,data=thisSite,id=GroupID,corstr="ar1")

JulianDay = c("Null","Linear","Smooth")
QICJ = c(QIC(mod0)[[1]],QIC(mod0a)[[1]],QIC(mod0b)[[1]])
rbind(JulianDay,QICJ)

# Normalized Time of Day
mod0c = geeglm(Presence~NormTime,family=binomial,data=thisSite,id=GroupID,corstr="ar1")
mod0d = geeglm(Presence~mSpline(NormTime,
                                knots=quantile(thisSite$NormTime,probs=c(0.333,0.666)),
                                Boundary.knots=c(-1,1)),
               family=binomial,data=thisSite,id=GroupID,corstr="ar1")
NormTime = c("Null","Linear","Smooth")
QICN = c(QIC(mod0)[[1]],QIC(mod0c)[[1]],QIC(mod0d)[[1]])
rbind(NormTime,QICN)

# Lunar Illumination
mod0e = geeglm(Presence~LunarIllum,family=binomial,data=thisSite,id=GroupID,corstr="ar1")
mod0f = geeglm(Presence~mSpline(LunarIllum,
                                knots=quantile(thisSite$LunarIllum,probs=c(0.333,0.666)),
                                Boundary.knots=c(0,1)),
               family=binomial,data=thisSite,id=GroupID,corstr="ar1")
Lunar = c("Null","Linear","Smooth")
QICL = c(QIC(mod0)[[1]],QIC(mod0e)[[1]],QIC(mod0f)[[1]])
rbind(Lunar,QICL)

````
## Covar Selection

Fit series of reduced models including all but one covar & use QIC to determine which covars to keep in model
``` {r covar selection}
# FullMod = geeglm(Presence~mSpline(JulianDay,
#                                       knots=quantile(thisSite$JulianDay, probs=c(0.333,0.666)),
#                                       Boundary.knots=c(1,365),
#                                       periodic=T)
#                      +mSpline(NormTime,
#                               knots=quantile(thisSite$NormTime,probs=c(0.25,0.5,0.75)),
#                               Boundary.knots=c(-1,1))
#                      +mSpline(LunarIllum,
#                               knots=)
#                      +as.factor(Year),
#                      +as.factor(DayPhase),
#                      +as.factor(HATSite),
#                      family=binomial,
#                      data=thisSite,
#                      id=reducedClustID,
#                      corstr="ar1")

CovSec1 = geeglm(Presence~mSpline(JulianDay,
                                      knots=quantile(thisSite$JulianDay, probs=c(0.333,0.666)),
                                      Boundary.knots=c(1,365),
                                      periodic=T)
                     +mSpline(LunarIllum,
                              knots=quantile(thisSite$LunarIllum,probs=c(0.333,0.666)),
                              Boundary.knots=c(0,1))
                     +as.factor(Year)
                     +as.factor(DayPhase),
                     family=binomial,
                     data=thisSite,
                     id=GroupID,
                     corstr="ar1")
CovSec1a = geeglm(Presence~mSpline(LunarIllum,
                              knots=quantile(thisSite$LunarIllum,probs=c(0.333,0.666)),
                              Boundary.knots=c(0,1))
                     +as.factor(Year)
                     +as.factor(DayPhase),
                     family=binomial,
                     data=thisSite,
                     id=GroupID,
                     corstr="ar1")
CovSec1b = geeglm(Presence~mSpline(JulianDay,
                                      knots=quantile(thisSite$JulianDay, probs=c(0.333,0.666)),
                                      Boundary.knots=c(1,365),
                                      periodic=T)
                     +as.factor(Year)
                     +as.factor(DayPhase),
                     family=binomial,
                     data=thisSite,
                     id=GroupID,
                     corstr="ar1")
CovSec1c = geeglm(Presence~mSpline(JulianDay,
                                      knots=quantile(thisSite$JulianDay, probs=c(0.333,0.666)),
                                      Boundary.knots=c(1,365),
                                      periodic=T)
                     +mSpline(LunarIllum,
                              knots=quantile(thisSite$LunarIllum,probs=c(0.333,0.666)),
                              Boundary.knots=c(0,1))
                     +as.factor(DayPhase),
                     family=binomial,
                     data=thisSite,
                     id=GroupID,
                     corstr="ar1")
CovSec1d = geeglm(Presence~mSpline(JulianDay,
                                      knots=quantile(thisSite$JulianDay, probs=c(0.333,0.666)),
                                      Boundary.knots=c(1,365),
                                      periodic=T)
                     +mSpline(LunarIllum,
                              knots=quantile(thisSite$LunarIllum,probs=c(0.333,0.666)),
                              Boundary.knots=c(0,1))
                     +as.factor(Year),
                     family=binomial,
                     data=thisSite,
                     id=GroupID,
                     corstr="ar1")
whichMiss = c("All In","MinusJDay","MinusLunIllum","MinusYear","MinusDayPhase")
QIC_CovSec1 = c(QIC(CovSec1)[[1]],QIC(CovSec1a)[[1]],QIC(CovSec1b)[[1]],QIC(CovSec1c)[[1]],QIC(CovSec1d)[[1]])
rbind(whichMiss,QIC_CovSec1)
```
### Construct Model

Covars enter model in order determined by QIC score from above (the ones that, if removed, determine the biggest increase in QIC, enter the model first)
Remove non-significant parameters from model

```{r covar order}
QIC_CovSec1[2:5] - QIC_CovSec1[1]
```

``` {r model construction}
TryMod1 = geeglm(Presence~as.factor(Year)
                      +mSpline(JulianDay,
                                      knots=quantile(thisSite$JulianDay, probs=c(0.333,0.666)),
                                      Boundary.knots=c(1,365),
                                      periodic=T)
                     +mSpline(LunarIllum,
                              knots=quantile(thisSite$LunarIllum,probs=c(0.333,0.666)),
                              Boundary.knots=c(0,1))
                     +as.factor(DayPhase),
                     family=binomial,
                     data=thisSite,
                     id=GroupID,
                     corstr="ar1")
signif = getPvalues(TryMod1)
signif

```
### Final Model

```{r final model}

FinalMod = geeglm(Presence~mSpline(JulianDay,
                                      knots=quantile(thisSite$JulianDay, probs=c(0.333,0.666)),
                                      Boundary.knots=c(1,365),
                                      periodic=T)
                     +mSpline(LunarIllum,
                              knots=quantile(thisSite$LunarIllum,probs=c(0.333,0.666)),
                              Boundary.knots=c(0,1))
                     +as.factor(Year)
                     +as.factor(DayPhase),
                     family=binomial,
                     data=thisSite,
                     id=GroupID,
                     corstr="ar1")

  sinkName = paste(seasDir,'/',CTname,'/',sites[j],"_",int,"_GEEGLMSummary.txt",sep="")
  sink(sinkName)
  getPvalues(FinalMod)
  sink()
```
### Bootstrap CIs & Plot


``` {r plot}

# Bootstrap GEEGLM parameter estimates for later construction of confidence intervals
  BootstrapParameters<-rmvnorm(10000, coef(modJday), summary(modJday)$cov.unscaled)
  JDayBootstrapCoefs<- BootstrapParameters[,2:3]
  YearBootstrapCoefs<- BootstrapParameters[,c(1,4:6)]
  if (j==7){ # if HAT, get coefficients for Site
    hatSiteBootstrapCoefs<- BootstrapParameters[,c(1,7)]
    quant.func<- function(x){quantile(x, probs=c(0.025,0.975))}
    cisH<-apply(hatSiteBootstrapCoefs, 2, quant.func)
  }
  
  # Predict presence at each X value using model coefficients
  if (j==7){ # if HAT, include Site as factor
    Jx1<-model.matrix(modJday)[,2:3]%*%coef(modJday)[c(2:3)]
    Jx2<-model.matrix(modJday)[,c(1,4:6)]%*%coef(modJday)[c(1,4:6)]
    Jx3<-model.matrix(modJday)[,c(1,7)]%*%coef(modJday)[c(1,7)]
  } else {
    Jx1 = model.matrix(modJday)[,2:3]%*%coef(modJday)[c(2:3)]
    Jx2 = model.matrix(modJday)[,c(1,4:6)]%*%coef(modJday)[c(1,4:6)]
  }
  
  
  # Plot GEEGLM partial residuals
  # Julian Day
  JDayForPlotting<- seq(min(Jday), max(Jday), length=256)
  Basis<- mSpline(JDayForPlotting,  # spline spanning range of X values
                  knots=quantile(Jday, probs=c(0.333,0.666)),
                  Boundary.knots=c(1,365),
                  periodic=T) # basis functions for smooth function
  RealFit<- Basis%*%coef(modJday)[c(2:3)] # multiply basis functions by model coefficients to get values of spline at each X
  RealFitCenterJ<- RealFit-mean(Jx1)-coef(modJday)[1] # adjust offset
  JDayBootstrapFits<- Basis%*%t(JDayBootstrapCoefs) # get spread of spline values at each X based on distributions of each coefficient
  quant.func<- function(x){quantile(x, probs=c(0.025,0.975))}
  cisJ<-apply(JDayBootstrapFits, 1, quant.func) # confidence interval of smooth function estimate
  Jcil<-cisJ[1,]-mean(Jx1)-coef(modJday)[1] # lowerCI bound
  Jciu<-cisJ[2,]-mean(Jx1)-coef(modJday)[1] # upper CI bound
  
  
  plotDF = data.frame(JDayForPlotting,RealFitCenterJ)
  colnames(plotDF) = c("Jday","Fit")
  dJday = stats::density(Jday,na.rm = TRUE,n=256,from=1,to=365) # Calculate kernel density of Jday observations
  dens = data.frame(c(dJday$x,rev(dJday$x)),c(dJday$y,rep(0,length(dJday$y))))
  colnames(dens) = c("Day","Density")
  dens$Density = dens$Density/max(dens$Density) # normalize kernel density
  if (min(Jcil)<0){ # set kernel density at bottom of y axis
    dens$Density = dens$Density - abs(min(Jcil)) 
  } else {
    dens$Density = dens$Density + min(Jcil)
  }
  
  #saveName = paste(seasDir,'/',CTname,'/',sites[j],"_",int,"_GEEGLM_JDayPlot.png",sep="")
  JD = ggplot(plotDF, aes(Jday, Fit),
  ) + geom_polygon(data=dens,
                   aes(Day,Density),
                   fill=4,
                   alpha=0.2
  ) + geom_smooth(fill = "grey",
                  colour = "black",
                  aes(ymin=Jcil, ymax=Jciu),
                  stat ="identity"
  ) + labs(x = "Julian Day",
           y = "s(Julian Day)",
           title = paste(CTname, 'at',sites[j]),
  ) + theme(axis.line = element_line(),
            panel.background = element_blank()
  )
  
  ggsave(saveName,device="png")
  while (dev.cur()>1) {dev.off()}
  
  
  # Year (as boxplot)
  # Center intercept (1st level of year factor) at 0 and show other levels relative to it
  AdjustedYearCoefs = data.frame(c(YearBootstrapCoefs[,1]-mean(YearBootstrapCoefs[,1]),
                                   YearBootstrapCoefs[,2],
                                   YearBootstrapCoefs[,3],
                                   YearBootstrapCoefs[,4]),
                                 as.factor(rep(2016:2019,each=10000)))
  colnames(AdjustedYearCoefs) = c("Coefficient","Year")
  # AdjustedYearCoefs = apply(AdjustedYearCoefs,2,exp) return to units of response var?
  
  #saveName = paste(seasDir,'/',CTname,'/',sites[j],"_",int,"_GEEGLM_YearPlot.png",sep="")
  Yr = ggplot(AdjustedYearCoefs,aes(Year,Coefficient)
  ) + geom_boxplot(
  )+ theme(axis.line = element_line(),
           panel.background = element_blank()
  ) + labs(title = paste(CTname, 'at',sites[j]))
  ggsave(saveName,device="png")
  while (dev.cur()>1) {dev.off()}

```

